Manuscript Draft

### 2019-01-11 manuscript.tex

\usepackage{lineno}

@@

\textbf{Describe why we compare all pairwise combos rather than
filtering for just the drug-gene combos in G2P}

Using cell-line-specific point mutation annotations from the CCLE, we
compared z-score-transformed area under the dose-response curve (AUC)
distributions between mutant and wildtype cell lines for all pairwise
combinations of G2P genes and drugs tested in the CCLE, CTRP, and GDSC
drug screens (\emph{Figures 1B and 2A}). At an unadjusted
\(\alpha = 0.05\), Wilcoxon rank-sum tests revealed that in CCLE, AUC
significantly correlated with mutation status for 18 out of 170
(10.59\%) tested drug-gene combinations were significantly different,
135 out of 918 (14.71\%) were significantly different for CTRP, and 89
out of 850 (10.47\%) for GDSC.

Controlling for a false discovery rate (FDR) \textless{} 5\%,
selumetinib-\emph{BRAF} (\(p_{adj} = 1.4^{-10}\)) was the only
significant drug-gene combination tested in CCLE (\(p_k = 7.9^{-4}\)).
CTRP (\(p_k = 6.5^{-4}\)) yielded 5 significant combinations, one of
which was selumetinib-\emph{BRAF}. There were 3 significant G2P
associations in the GDSC dataset (\(p_k = 9.5^{-5}\)), Of all the
significant correlations between mutation status and AUC across
datasets, the only genes represented were \emph{KRAS} and \emph{BRAF}
(Figure~\ref{fig:drug-venn}).

\begin{verbatim}
## NULL
\end{verbatim}

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/drug-venn-1.pdf}
\caption{}
\end{figure}

We observe little concordance between the number of interpretations for
each G2P association and the significance of mutation status on AUC for
said G2P association, suggesting that mutation status alone is not a
robust predictor of clinically actionable genetic targets. This led us
to examine the correlations between AUC and gene expression and AUC and
copy number, two other genomic features that are known to associate with
cancer and drug response.

We computed Spearman correlation coefficients (\(r_s\)) between AUC of
each drug-gene association and gene expression (RPKM) and copy number
(log\(_2\) ratio) of the gene in the association\ldots{}

\textbf{Cut this analysis since we are comparing distributions of
p-values?}

@@

In the level-A G2P interpretations, there were 102 unique
drug-gene-lineage combinations (DGLs), representing seven of our curated
cancer lineages: cancer, lung cancer, breast cancer, leukemia, thyroid
cancer, skin cancer, and ovarian cancer. CCLE had data for 9 DGLs, CTRP
had data for 61, and GDSC had data for 53, yielding a total of 73 unique
DGLs across all three datasets. Only 7 DGLs were screened in all three
datasets,

@@

@@

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/crispr-venn-1.pdf}
\caption{}
\end{figure}

@@

For the purpose of comparison to the drug screen data, if we treat
CRISPR knockdown like a small-molecule drug, then these results are
similar to the results from CCLE, CTRP, and GDSC (Table 2).

@@

Wilcoxon tests were conducted using the \texttt{compare\_means} function
in the ggpubr (v0.1.7.999) R package (Kassambara, 2018).
\texttt{compare\_means} does Benjamini-Hochberg p-value adjustment using
the \texttt{p,adjust} function from the \texttt{stats} base R package.

For computing Spearman correlation coefficients. If there were 3 or
fewer data points, then the correlation was not done and a value of NA
was supplied.

>>>>

\DeclareDelayedFloatFlavor{ThreePartTable}{table}
\DeclareDelayedFloatFlavor{lltable}{table}
\DeclareDelayedFloatFlavor*{longtable}{table}
\makeatletter
\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
\makeatother
\usepackage{lineno}

@@

Using cell-line-specific point mutation annotations from the CCLE, we
compared z-score-transformed area under the dose-response curve (AUC)
distributions between mutant and wildtype cell lines for all G2P
associations tested in the CCLE, CTRP, and GDSC drug screens. Of the 88
G2P associations, data was available for 9 associations in CCLE, 48 in
CTRP, and 44 in GDSC. Controlling for a false discovery rate (FDR)
\textless{} 5\%, CCLE yielded no significant correlations (Wilcoxon
rank-sum tests, \(p_k = 2.1^{-4}\)), though the most significant G2P
association was selumetinib-\emph{KRAS} (\(p_{adj} = 1.9^{-3}\)). In
CTRP, AUC from 3 drug-gene associations correlated with the gene's
mutation status (\(p_k = 4.5^{-3}\)): selumetinib-\emph{KRAS}
(\(p_{adj} = 4.4^{-8}\)), vemurafenib-\emph{BRAF}
(\(p_{adj} = 4.4^{-8}\)), and dabrafenib-\emph{BRAF}
(\(p_{adj} = 7.7^{-5}\)). Dabrafenib-\emph{BRAF}
(\(p_{adj} = 8.6^{-11}\)) and trametinib-\emph{BRAF}
(\(p_{adj} = 4.2^{-6}\)) were significant in GDSC (\(p_k = 2.3^{-3}\)).

\begin{quote}
\textbf{(???) Describe why we compare all pairwise combos rather than
filtering for just the drug-gene combos in G2P?}
\end{quote}

To test for potentially novel drug-gene associations, we conducted
similar Wilcoxon rank-sum tests for all pairwise combinations of G2P
genes and drugs tested in the three drug screens (\emph{Figures 1B and
2A}). At an unadjusted \(\alpha = 0.05\), the tests revealed that in
CCLE, AUC significantly correlated with mutation status for 18 out of
170 (10.59\%) tested drug-gene combinations were significantly
different, 135 out of 918 (14.71\%) were significantly different for
CTRP, and 89 out of 850 (10.47\%) for GDSC.

Controlling for a FDR \textless{} 5\%, selumetinib-\emph{BRAF}
(\(p_{adj} = 1.4^{-10}\)) was the only significant drug-gene combination
tested in CCLE (\(p_k = 7.9^{-4}\)). CTRP (\(p_k = 6.5^{-4}\)) yielded 5
significant combinations, one of which was selumetinib-\emph{BRAF}.
There were 3 significant G2P associations in the GDSC dataset
(\(p_k = 9.5^{-5}\)). Of all the significant correlations between
mutation status and AUC across datasets, the only genes represented were
\emph{KRAS} and \emph{BRAF} (Figure~\ref{fig:drug-venn}). Of the 7
significant drug-gene associations across datasets, 4 were G2P
associations. The remaining 3 associations were selumetinib-\emph{BRAF},
trametinib-\emph{KRAS}, and pazopanib-\emph{BRAF}.

\begin{quote}
\textbf{(???) Trametinib and selumetinib are both \emph{MEK} inhibitors,
and pazopanib is a \emph{VEGFR} inhibitor. VEGF signaling can influence
the MAPK/ERK signaling pathway, which promotes cell proliferation and
survival. This pathway includes \emph{RAS}, \emph{BRAF}, and \emph{MEK}
proteins. These results underscore the importance of the MAPK/ERK
signaling pathway in cancer.}
\end{quote}

However, taking these results as a whole, we observe little concordance
between G2P associations and the correlation of mutation status and AUC
for said G2P association, suggesting that mutation status alone is not a
robust predictor of clinically actionable genetic targets. This led us
to examine the correlations between AUC and gene expression and AUC and
copy number, two other genomic features that are known to associate with
cancer and drug response. We computed Spearman correlation coefficients
(\(r_s\)) between AUC of each drug-gene association and gene expression
(RPKM) and copy number (log\(_2\) ratio) of the gene in the
association\ldots{}

\begin{figure}
\centering
\includegraphics{./plots/manuscript/drugscreen_venn.png}
\caption{}
\end{figure}

\begin{quote}
\textbf{(???) Cut this analysis since we are comparing distributions of
p-values?}
\end{quote}

@@

In the level-A G2P interpretations, there were 102 unique
drug-gene-lineage combinations (DGLs), representing seven of our curated
cancer lineages: cancer, lung cancer, breast cancer, leukemia, thyroid
cancer, skin cancer, and ovarian cancer. CCLE had data for 9 DGLs, CTRP
had data for 61, and GDSC had data for 53, yielding a total of 73 unique
DGLs across all three datasets. Only 7 DGLs were screened in all three
datasets.

@@

\begin{quote}
\textbf{(???) KS test results}
\end{quote}

(Figure~\ref{fig:ksplot})

\begin{figure}
\centering
\includegraphics{./plots/manuscript/g2p_spearman_ks_res_plot.png}
\caption{\label{fig:ksplot}Comparison of Spearman correlation coefficients
derived from copy number (log\(_2\) ratio), gene expression (RPKM), and
AUC z-scores from G2P drug-gene associations. Coefficients were grouped
by whether or not the associated lineage for each comparison was in a
G2P drug-gene-lineage association or not. Distributions of these grouped
coefficients were compared using Kolmogorov-Smirnov tests.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{./plots/manuscript/crispr_venn.png}
\caption{}
\end{figure}

@@

For the purpose of comparison to the drug screen data, if we treat
CRISPR knockdown like a small-molecule drug, then these results are more
comprehensive than the results from CCLE, CTRP, and GDSC (Table 2).

@@

Wilcoxon tests were conducted using the \texttt{compare\_means} function
in the ggpubr (v0.1.7.999) R package (Kassambara, 2018).
\texttt{compare\_means} does Benjamini-Hochberg p-value adjustment using
the \texttt{p.adjust} function from the \texttt{stats} base R package.

For computing Spearman correlation coefficients, we used the
\texttt{cor.test} base R package. If there were 3 or fewer data points,
then the correlation was not done and a value of \texttt{NA} was
supplied.
